Inferring the effective start dates of non-pharmaceutical interventions during \covid outbreaks

### 2020-08-26 ms/ms.tex
prep for ploscompbiol

@@

@@

Our results highlight the complex interaction between personal, regional, and global determinants of behavioral response to an epidemic.
Therefore, we emphasize the need to further study variability in compliance and behavior over both time and space. This can be accomplished both by surveying differences in compliance within and between populations~\citep{Atchison2020}, and by incorporating specific behavioral models into epidemiological models~\citep{Arthur2020,Fenichela2011,Walters2013}.

@@

%\pagebreak
% Acknowledgements
{\small
\section*{Acknowledgements}
We thank Yinon M. Bar-On, Lilach Hadany, and Oren Kolodny for discussions and comments.
This work was supported in part by the Israel Science Foundation (3811/19, 552/19, and 1399/17).
}

>>>>

Keywords: SEIR, COVID-19, public health, pandemic, infectious disease, NPI

%Classification: Biological Sciences / Population Biology; Social Sciences / Environmental Sciences

%\pagebreak

@@

%\renewcommand{\abstractname}{Significance statement}
%\begin{abstract}
%We have inferred the effective start date of non-pharmaceutical interventions, like lockdowns, during the initial \covid outbreak by fitting an epidemiological model to daily case data in 12 countries. Our results show that the effective start of interventions in many cases occurs 3-10 days later then the official day. Moreover, we demonstrate that such delays can lead to under-estimation of the intervention impact, and thus cause decision makers to implement stricter interventions. We suggest that a late effective start of interventions may be the result of low compliance to authority or conformity to social norms. Our study therefore highlights the importance of further research into the role of human behavior and social organization on the current and future epidemics.
%\end{abstract}

@@

Our results highlight the complex interaction between personal, regional, and global determinants of behavioral response to an epidemic.
Therefore, we emphasize the need to further study heterogeneity in compliance and behavior over both time and space. This can be accomplished both by surveying differences in compliance within and between populations~\citep{Atchison2020}, and by incorporating specific behavioral models into epidemiological models~\citep{Arthur2020,Fenichela2011,Walters2013}.

@@

%\pagebreak
% Acknowledgements
{\small
\section*{Acknowledgements}
We thank Yinon M. Bar-On, Lilach Hadany, and Oren Kolodny for discussions and comments.
This work was supported in part by the Israel Science Foundation (3811/19 and 552/19, YR).
}
